---
title: "Homework 3"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r load data}
data("instacart")
```

This dataset has `r nrow(instacart)` rows and `r ncol(instacart)` columns.

There are variables based on the user/order -- user ID, order ID, order number, order day, order hour. There are also variables based on the item -- name, aisle, department, product ID.

#### How many aisles, and which aisles are the most items ordered from?

```{r aisles}
aisles_count = 
  instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

There are `r nrow(aisles_count)` aisles and most items are ordered from fresh vegetables and fresh fruits.

#### Time to make a plot

```{r aisles plot}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>% 
  ggplot(aes(x = aisle, y = n))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

#### Three most popular items in the aisles "baking ingredients", "dog food care", and "packaged vegetables fruits".

```{r table popular items}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

#### Table for mean hour of the day where Pink Lady Apples and Coffee Ice Cream are ordered

```{r apples n ice cream}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```

## Problem 2

```{r import data}
read.csv("data/accel_data.csv") %>% 
  pivot_longer( activity.1:activity.1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity_count"
  ) %>% 
  mutate(
    day = factor(day),
    minute = as.numeric(minute),
    weekday_weekend = recode(day, "Monday" = "weekday", "Tuesday" = "weekday", "Wednesday" = "weekday", "Thursday" = "weekday", "Friday" = "weekday", "Saturday" = "weekend", "Sunday" = "weekend"))
```











